---
title: "Documentation in R"
author: "Rebecca Barter"
date: "2/24/2017"
output: html_document
---

## Caret: an example of good documentation

### Load in package
```{r eval = FALSE} 
install.packages("lubridate")
```

```{r message=FALSE}
library(lubridate)
```


### Look at the help files

```{r}
?ymd
```

### Look at the vignette

```{r eval = F}
vignette("lubridate")
```


## Create our own package and document it

If you are going to be making your own package, you will need to install `devtools`.

```{r eval = FALSE}
install.packages("devtools")
```

### Create package skeleton

Then we can create a new package skeleton that devtools will fill out with the basic files.

```{r eval = FALSE}
devtools::create("path/to/package/pkgname")
```


In this package skeleton, you will find a folder, `R`. This is where your functions will go.

### Add some functions

Inside the `R` folder, we will add the files `add.R` and `multiply.R` which will contain functions which can add and multiply two numbers together.

For example, `add.R` will contain the following function:

```{r}
add <- function(x, y) {
  x + y
}
```

and `multiply.R` will contain the following function:

```{r}
multiply <- function(x, y) {
  x * y
}
```


### Document these functions

We can now use roxygen comments to document these functions. This will mean that when we type `?add`, a help files will appear!

Roxygen comments, #', look a lot like regular R comments, #. To document the `add()` function, we will add comments above the function in the `add.R` file which will include tags for parameters (`@param`), the return value (`@return`) and examples (`@examples`). The first sentence is the title of the document.

```{r eval = FALSE}
#' Take the sum of two values
#'
#' @param x a numeric value
#' @param y a numeric value
#' 
#' @return the sum of x and y
#' 
#' @examples 
#' add(3, 4)
#' add(4, 3)

add <- function(x, y) {
  x + y
}

```

We can roxygenize the multiply function in a similar manner.
```{r eval = F}
#' Take the scalar product of two values
#'
#' @param x a numeric value
#' @param y a numeric value
#' 
#' @return the product of x and y
#' 
#' @examples 
#' multiply(3, 4)
#' multiply(4, 3)

multiply <- function(x, y) {
  x * y
}
```

### Render the help files

You can render the documentation for your package using the `document()` function from the devtools package. Make sure that you are in the package working directory!

```{r eval = F}
devtools::document()
```

Now if you type `?add`, the help file should appear in the help panel.

How easy is that!


### Building a vignette

Next, we can build a vignette. If you want a pdf vignette, you can create a new "R Sweave" file, or to create a html vignette, you can create a new "R Markdown" file. The vignette will go in a folder appropriately titled `vignette/`. 

Unsurprisingly, devtools offers a function for creating a new vignette. Let's create a new vignette called "math"

```{r eval = FALSE}
devtools::use_vignette("math")
```


A file called `math.Rmd` file should have opened up. Knit it to see an example of the output.

We can edit this vignette to describe how to use our package.


### Bookdown documentation

If your package is complex and has many different types of functionalities, a single vignette will likely not be enough. While you can certainly make multiple vignettes, a better way is to write a documentation "book" (techincally it is a gitbook) using bookdown! Although this isn't necessarily an official form of documentation for R packages yet, it is becoming more and more common. 

To create a gitbook, you need to clone the demo from github. To do so, in your terminal, type the following git command

```{r engine = 'bash', eval = F}
git clone https://github.com/rstudio/bookdown-demo
```

A folder names `bookdown-demo/` will have appeared in your working directory folder. Navigate to it.

```{r, eval = F}
setwd("path/to/bookdown-demo")
```

Next, install the bookdown package.

```{r eval = F}
install.packages("bookdown")
```

You can compile the book template using the `render_book()` command. Be sure to include the argument `output_format = bookdown::gitbook()` if you want to compile a gitbook (rather than a .pdf).

```{r eval = F}
bookdown::render_book("index.Rmd", output_format = bookdown::gitbook())
```

A `_book/` folder will have appeared in your bookdown-demo folder. In this folder, open the `index.html` file.
